---
title: "Lab | Exploring the sample_superstore Dataset"
output: html_notebook
editor_options: 
  markdown: 
    wrap: sentence
---

# Lab \| Exploring the sample_superstore Dataset

In this lab, you will work with the sample_superstore dataset to practice creating, inspecting, and manipulating dataframes.

Follow the steps below to complete the tasks.

## Step 1: Importing the Dataset Load the sample_superstore dataset into a dataframe.

You can find this dataset in the datasets package or download it as a CSV file from online sources.
Save the dataframe as superstore.

```{r}
# Importing a CSV file
df <- read.csv("Sample - Superstore.csv")
library(dplyr)


```

Hint: Use the read.csv() or read_excel() function to import the dataset.

## Step 2: Inspecting the Dataframe Display the first 10 rows of the dataframe using the head() function.

Use the str() function to inspect the structure of the dataframe.

```{r}
str(df)
```

-   What are the data types of the columns?
    -   answ: int, chr, num
-   Use the summary() function to get a summary of the dataframe. What insights can you gather from the summary?

```{r}
summary(df)
```

## Step 3: Accessing Data Extract the Sales column as a vector using the \$ operator.

```{r}
sales_vector <- df$Sales
```

Subset the first 15 rows and the columns Order ID, Customer Name, and Sales.
Use the nrow() and ncol() functions to determine the number of rows and columns in the dataframe.

```{r}
# Extract the Sales column as a vector
sales_vector <- df$Sales

# Subset the first 15 rows and specific columns
subset_data <- df[1:15, c("Order.ID", "Customer.Name", "Sales")]

# Determine the number of rows and columns
number_of_rows <- nrow(df)
number_of_columns <- ncol(df)

```

## Step 4:

-   Filtering Data Filter the dataframe to show only rows where the Profit is greater than 100.

    ```{r}
    profit_over_100 <- df[df$Profit > 100, ]
    ```

-   Filter the dataframe to show only rows where the Category is "Furniture" and the Sales are greater than 500.

```{r}
furniture_sales_over_500 <- df[df$Category == "Furniture" & df$Sales > 500, ]
```

-   Filter the dataframe to show only rows where the Region is "West" and the Quantity is greater than 5.

```{r}

west_quantity_over_5 <- df[df$Region == "West" & df$Quantity > 5,]
```

## Step 5:

-   Adding and Modifying Columns Add a new column called Profit Margin that calculates the profit margin as (Profit / Sales) \* 100.

```{r}

# Use transform() to add a new column and modify an existing one
df <- transform(df,
                Profit_Margin = (Profit / Sales) * 100  # Add Profit Margin column
                #round(Sales, 2)                  # Modify Sales column
)



```

-   Modify the Sales column by rounding the values to 2 decimal places.

```{r}
df <- transform(df, Sales = round(Sales, 2)                  # Modify Sales column
)

```

-   Remove the Postal Code column from the dataframe using the subset() or select() function.

```{r}
# Remove the Postal Code column using select()
df <- select(df, -Postal.Code)
```

## Step 6:

-   Handling Missing Data Check for missing values in the dataframe using the is.na() function.

```{r}
# Check for missing values
sum(is.na(df))

# See if there are any missing values in the data frame
any_missing <- any(missing_values)
print(any_missing)
```

-   Are there any missing values?

    -   no

-   If there are missing values, remove rows with missing data using the na.omit() function.

```{r}
na.omit(df)
```

-   Replace any missing values in the Sales column with the mean of the Sales column using the na.fill() function.

```{r}
# Calculate the mean of 'Sales', excluding NA values
sales_mean <- mean(df$Sales, na.rm = TRUE)

# Replace NA values in the 'Sales' column with the mean
df$Sales[is.na(df$Sales)] <- sales_mean
```

## Step 7: Advanced Analysis (Optional)

-   Group the dataframe by Region and calculate the total Sales and Profit for each region.
```{r}
# Load dplyr for data manipulation
library(dplyr)

# Group by Region and calculate total Sales and Profit
region_summary <- df %>%
  group_by(Region) %>%
  summarize(Total_Sales = sum(Sales, na.rm = TRUE),
            Total_Profit = sum(Profit, na.rm = TRUE))

```

-   Create a new column called Discount Level that categorizes the Discount column into: - "Low" (0-0.2)
- "Medium" (0.2-0.5)
- "High" (0.5-1)

```{r}
# Categorize the Discount column
df$Discount_Level <- cut(df$Discount,
  breaks = c(-Inf, 0.2, 0.5, 1),
  labels = c("Low", "Medium", "High"),
  right = FALSE)

```



- Sort the dataframe by Sales in descending order.

```{r}
# Sort the dataframe by Sales in descending order
df <- df %>%
  arrange(desc(Sales))
```

